```         
---
title: "Build and deploy a stroke prediction model using R"
date: "`r Sys.Date()`"
output: html_document
author: "Dhruvi Swadia"
---
```

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics and building the prediction models. The final report was completed on `r date()`.

**Data Description:**

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.

```{r}
if(!require('tidyverse')){
  install.packages('tidyverse')
  library('tidyverse')}
```

```{r}
if(!require('lubridate')){
  install.packages('lubridate')
  library('lubridate')
}
if(!require('ggplot2')){
  install.packages('ggplot2')
  library('ggplot2')
}
if(!require('dplyr')){
  install.packages('dplyr')
  library('dplyr')
}
if(!require('caret')) {
  install.packages('caret')
  library('caret')
}
if(!require('randomForest')) {
  install.packages('randomForest')
  library('randomForest')
}
if(!require('skimr')) {
  install.packages('skimr')
  library('skimr')
}
if(!require('gridExtra')) {
  install.packages('gridExtra')
  library('gridExtra')
}
if(!require('caTools')) {
  install.packages('caTools')
  library('caTools')
}

```

```{r}
if(!require('corrplot')) {
  install.packages('corrplot')
  library('corrplot')
}
if(!require('ggcorrplot')) {
  install.packages('ggcorrplot')
  library('ggcorrplot')
}
## Loading required package: ggcorrplot
if(!require('naniar')){
  install.packages('naniar')
  library('naniar')
}
```

```{r}
#importing files
setwd("/Users/dhruviswadia/Desktop/Projects/Stroke prediction-R")
Data_Stroke <- read.csv('healthcare-dataset-stroke-data.csv')
```

```{r}
summary(Data_Stroke)
```

```{r}
glimpse(Data_Stroke)
```

```{r}
skim(Data_Stroke)
```

```{r}
miss_scan_count(data = Data_Stroke, search = list("Unknown","N/A","Other"))
```

```{r}
##Convert NA to median in BMI column
Data_Stroke$bmi <- as.numeric(Data_Stroke$bmi)
```

```{r}
idx <- complete.cases(Data_Stroke)
bmi_idx <- is.na(Data_Stroke$bmi)
median_bmi <- median(Data_Stroke$bmi, na.rm = TRUE)

Data_Stroke[bmi_idx,]$bmi <- median_bmi
colSums(is.na(Data_Stroke))
```

```{r}
##Check duplicates
sum(duplicated(Data_Stroke))
```

```{r}
Data_Stroke %>% count(gender)
```

```{r}
##Remove ID and filter out 'Other' values in Gender
Data_Stroke <- Data_Stroke %>% 
  select(-c(id)) %>% 
  filter(gender != "Other")
str(Data_Stroke)
```

```{r}
##Convert non-numeric variables to factors
Data_Stroke$stroke <- factor(Data_Stroke$stroke, levels = c(0,1), labels = c("No", "Yes"))
Data_Stroke$hypertension <- factor(Data_Stroke$hypertension, levels = c(0,1), labels = c("No", "Yes"))
Data_Stroke$heart_disease <- factor(Data_Stroke$heart_disease, levels = c(0,1), labels = c("No", "Yes"))
```

```{r}
#Task 2
```

```{r}
d1 <- Data_Stroke %>%
  ggplot(aes(x = gender, fill = gender)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Gender Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)
  
d2 <- Data_Stroke %>%
  ggplot(aes(x = hypertension, fill = hypertension)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Hypertenstion Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)
  

d3 <- Data_Stroke %>%
  ggplot(aes(x = heart_disease, fill = heart_disease)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Heart Disease Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d4 <- Data_Stroke %>%
  ggplot(aes(x = ever_married, fill = ever_married)) +
  geom_bar(fill = c("red","blue")) +
  ggtitle("Married distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d5 <- Data_Stroke %>%
  ggplot(aes(x = work_type, fill = work_type)) +
  geom_bar(fill = c("red", "blue","green","orange","aquamarine")) +
  ggtitle("Work type distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d6 <- Data_Stroke %>%
  ggplot(aes(x = stroke, fill = stroke)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Stroke distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d7 <- Data_Stroke %>%
  ggplot(aes(x = Residence_type, fill = Residence_type)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Residence distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)


grid.arrange(d1,d2,d3,d4,d5,d6,d7, ncol=2)
```

```{r}
#different visualizations 
#Plot Gender Distribution
Data_Stroke %>%
  ggplot(aes(x = gender, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "#FF7FAA")) +
  ggtitle("Gender vs. Stroke")
```

```{r}
#Plot Residence Type Distribution
Data_Stroke %>%
  ggplot(aes(x = Residence_type, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("#8EE5EE",
                             "#013220")) +
  ggtitle("Residence type vs. Stroke")
```

```{r}
#Plot Smoking Status Distribution
Data_Stroke %>%
  ggplot(aes(x = smoking_status, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("brown3",
                             "#5BCBE2")) +
  ggtitle("Smoking status vs. Stroke")
```

```{r}
#Plot average glucose level: This R code creates a density plot to visualize the distribution of average glucose levels in the dataset, distinguishing between people who had a stroke and those who did not.

  
Data_Stroke %>%
  ggplot(aes(x = avg_glucose_level, fill = stroke)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet"
  )) +
  ggtitle("Average Glucose level vs. Stroke") 
```

```{r}
#compare BMI distributions between stroke and non-stroke groups
Data_Stroke %>% filter(between(bmi, 0, 60)) %>%
  ggplot(aes(x = bmi, fill = stroke)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values=c("aquamarine3",
                             "#800020"
  )) +
  ggtitle("Body Mass Index vs. Stroke")
```

```{r}
# Task Three: Evaluate and select prediction models
```

```{r}

sample.split(Data_Stroke$stroke,SplitRatio = 0.8)->split_tag
train<-subset(Data_Stroke,split_tag==TRUE)
test<-subset(Data_Stroke,split_tag==FALSE)
dim(train)
```

```{r}
dim(test)
```

```{r}
#Task 4 Deploy the prediction model

```

```{r}
set.seed(123)
rf <- randomForest(formula = stroke~.,data = train)
rf
```

```{r}
plot(rf)
```

```{r}
# Generate predictions using the random forest model on the test set
predictions <- predict(rf, test)

# Compute the confusion matrix
conf_matrix <- confusionMatrix(predictions, test$stroke)
print(conf_matrix)
```

```{r}
# Convert the confusion matrix to a data frame for ggplot
conf_matrix_df <- as.data.frame(as.table(conf_matrix))
colnames(conf_matrix_df) <- c("Predicted", "Actual", "Count")
# Check the content of the confusion matrix to ensure it has the right data
print(head(conf_matrix_df))  # Debugging step

```

```{r}
# Create a custom color map for each combination of Actual vs Predicted using "Yes" and "No"
color_map <- c("No.No" = "violet",   # Actual=No, Predicted=No
               "No.Yes" = "orange",  # Actual=No, Predicted=Yes
               "Yes.No" = "green",   # Actual=Yes, Predicted=No
               "Yes.Yes" = "blue")   # Actual=Yes, Predicted=Yes

# Combine Actual and Predicted into a new column for discrete color mapping
conf_matrix_df$ColorKey <- paste(conf_matrix_df$Actual, conf_matrix_df$Predicted, sep = ".")

# Convert Actual and Predicted to "Yes" and "No" for better readability
conf_matrix_df$Actual <- factor(conf_matrix_df$Actual, levels = c(0, 1), labels = c("No", "Yes"))
conf_matrix_df$Predicted <- factor(conf_matrix_df$Predicted, levels = c(0, 1), labels = c("No", "Yes"))

# Debugging: Check the unique values of ColorKey to confirm
print(unique(conf_matrix_df$ColorKey))  # Debugging step

# Convert ColorKey to a factor with levels in the same order as in color_map
conf_matrix_df$ColorKey <- factor(conf_matrix_df$ColorKey, levels = names(color_map))

# Debugging: Check the levels of ColorKey after conversion
print(levels(conf_matrix_df$ColorKey))  # Debugging step

# Plot the confusion matrix using ggplot2 with 4 different colors
ggplot(conf_matrix_df, aes(x = Actual, y = Predicted, fill = ColorKey)) +
  geom_tile() +
  geom_text(aes(label = Count), vjust = 1) + 
  scale_fill_manual(values = color_map) +
  labs(title = "Confusion Matrix - Random Forest", x = "Actual", y = "Predicted") +
  theme_minimal()
```
